---
title: "entropart tutorial"
output: html_notebook
---

Camillo lab discussion.

## [entropart: An R Package to Measure and Partition Diversity](https://doi.org/10.18637/jss.v067.i088)

entropart is a package for R designed to estimate diversity based on HCDT entropy or similarity-based entropy. It allows calculating species-neutral, phylogenetic and functional entropy and diversity, partitioning them and correcting them for estimation bias.

```{r}
#install.packages("entropart")
library(entropart)
```


```{r}
(df <- data.frame(C1 = c(10, 10, 10, 10), C2 = c(0, 20, 35, 5), C3 = c(25, 15, 0, 2), row.names = c("sp1", "sp2", "sp3", "sp4")))

colSums(df) #total abundance in each community
rowSums(df) #abundance of each species in total across communities
sum(df) #total number of individuals
```

```{r}
w <- c(1,2,1) #Community weight to account for (e.g.) differences in effort, study sites, etc.
MC <- MetaCommunity(Abundances = df, Weights = w) #
```

Big help vs other community packages is that the first column can contain the species names (vs just matrix with rownames)

```{r}
MC
str(MC)
MC$Ps #probability of occurrence of each species in the meta-community
MC$Psi
```


```{r}
plot(MC)
```

```{r}
data("Paracou618")
summary(Paracou618.MC)
```

# It can also do phylogenetic and functional diversity
```{r}
Tsallis(Ps = Paracou618.MC$Ps, q = 1)  # Order 1 is the weighted geometric mean (see above), also HCDT
Shannon(Ps = Paracou618.MC$Ps)  # Notice that is the same!
```


# Sample coverage - probability for a species of the community to be observed in the actual sample.
Coverage(Ns = Paracou618.MC$Ns)

# Bias-corrected estimators - improve the estimation of entropy despite unobserved species
bcTsallis(Ns = Paracou618.MC$Ns, q = 1)

# Bias-corrected Shannon - This one is the one we should use!
bcShannon(Ns = Paracou618.MC$Ns)

# Effective numbers of species - Entropy converted into “true diversity” via Hill's numbers
expq(Simpson(Ps = Paracou618.MC$Ps), q = 2)
Diversity(Ps = Paracou618.MC$Ps, q = 2)

# Effective species number with bias correction
bcDiversity(Ns = Paracou618.MC$Ns, q = 2)

# Meta-community functions
# Alpha entropy
e <- AlphaEntropy(Paracou618.MC, q = 1)
summary(e)


# Diversity partitioning among alpha, beta and gamma diversity
p <- DivPart(q = 1, MC = Paracou618.MC, Biased = FALSE)
summary(p)

#  Diversity estimates for each component
de <- DivEst(q = 1, Paracou618.MC, Biased = FALSE, Correction = "Best", Simulations = 1000)
summary(de)
plot(de)
#  Diversity profile
dp <- DivProfile(seq(0, 2, 0.2), Paracou618.MC, Biased = FALSE)
summary(dp)
plot(dp)
